Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-06-11 paper.qmd
Updated - values with proper notations

To illustrate the method, we use $7\text{-}D$ simulated data, which we call the "nonlinear clusters". It is constructed by simulating two clusters, each consisting of $1000$ observations. The C-shaped cluster is generated from $\theta \sim U(-3\pi/2, 0)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the C), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$, $X_4 = \cos(\theta)$. The other cluster is from $X_1 \sim U(0, 2)$, $X_2 \sim U(0, 3)$, $\gamma \sim U(0, 0.5)$, $X_3 = -(X_1^3 + X_2) + \gamma$, and $X_4 \sim U(0, 2)$. We would consider $T=(X_1, X_2, X_3, X_4)$ to be the geometric structure (true model) that we hope to capture.

@@

To ensure that the grid covers the range of data values a buffer parameter ($q$) is set as a proportion of the range. By default,  $q=0.1$. The buffer should be extending a full hexagon width ($a_1$) and height ($a_2$) beyond the data, in all directions. The lower left position where the grid starts is defined as $(s_1, s_2)$, and corresponds to the centroid of the lowest left hexagon, $C_{1}^{(2)} = (c_{11}, c_{12})$. This must be smaller than the minimum data value. Because it is one buffer unit, $q$ below the minimum data values, $s_1 = -q$ and $s_2 = -qr_2$.

>>>>

To illustrate the method, we use $7\text{-}D$ simulated data, which we call the "nonlinear clusters". It is constructed by simulating two clusters, each consisting of $1000$ observations. The C-shaped cluster is generated from $\theta \sim U(\text{-}3\pi/2, 0)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the C), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$, $X_4 = \cos(\theta)$. The other cluster is from $X_1 \sim U(0, 2)$, $X_2 \sim U(0, 3)$, $\gamma \sim U(0, 0.5)$, $X_3 = \text{-}(X_1^3 + X_2) + \gamma$, and $X_4 \sim U(0, 2)$. We would consider $T=(X_1, X_2, X_3, X_4)$ to be the geometric structure (true model) that we hope to capture.

@@

To ensure that the grid covers the range of data values a buffer parameter ($q$) is set as a proportion of the range. By default,  $q=0.1$. The buffer should be extending a full hexagon width ($a_1$) and height ($a_2$) beyond the data, in all directions. The lower left position where the grid starts is defined as $(s_1, s_2)$, and corresponds to the centroid of the lowest left hexagon, $C_{1}^{(2)} = (c_{11}, c_{12})$. This must be smaller than the minimum data value. Because it is one buffer unit, $q$ below the minimum data values, $s_1 = \text{-}q$ and $s_2 = \text{-}qr_2$.
